---
title: "Heston No-Jump Simulation: Daily BNS Jump Test with and without Leverage"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# 1) Packages

```{r packages}
required_pkgs <- c("highfrequency", "data.table", "xts", "zoo")
missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, logical(1), quietly = TRUE)]

if (length(missing_pkgs) > 0L) {
  stop(
    "Missing required package(s): ",
    paste(missing_pkgs, collapse = ", "),
    ". Install them before rendering this Rmd."
  )
}

library(highfrequency)
library(data.table)
library(xts)
library(zoo)
```

# 2) Model and Experiment Design

We simulate the no-jump Heston-type stochastic volatility model:

$$dX_t = \sqrt{V_t}\,dW_t,\qquad
dV_t = \kappa(\theta - V_t)\,dt + \xi\sqrt{V_t}\,dB_t,$$

with instantaneous correlation
$$\mathrm{corr}(dW_t, dB_t) = \rho.$$

- `rho = 0`: no leverage effect.
- `rho < 0`: leverage effect.

There is **no jump component** in the data generating process, so the BNS jump test should reject near the nominal significance level if well-sized.

# 3) Simulation/Test Parameters

```{r parameters}
set.seed(20260602)

# Monte Carlo design (override via options() for quick smoke tests)
n_paths <- as.integer(getOption("heston_bns_n_paths", 200L))
n_days <- as.integer(getOption("heston_bns_n_days", 252L))
steps_per_day <- as.integer(getOption("heston_bns_steps_per_day", 288L))  # 5-minute over 24h
confidence_grid <- getOption("heston_bns_confidence_grid", c(0.90, 0.95, 0.99))

# Heston parameters
heston_par <- list(
  kappa = 5.0,
  theta = 0.04,
  xi = 0.60,
  v0 = 0.04
)

# Leverage scenarios
scenario_tbl <- data.table(
  scenario = c("No leverage", "Leverage"),
  rho = c(0.0, -0.7)
)

# Additional analysis A: leverage level grid
leverage_rho_grid <- as.numeric(getOption(
  "heston_bns_leverage_rho_grid",
  c(0.0, -0.2, -0.4, -0.6, -0.8, -0.95)
))
leverage_rho_grid <- sort(unique(leverage_rho_grid), decreasing = TRUE)
n_paths_leverage_grid <- as.integer(getOption("heston_bns_n_paths_leverage_grid", 120L))
n_days_leverage_grid <- as.integer(getOption("heston_bns_n_days_leverage_grid", n_days))

# Additional analysis B: sampling interval grid at fixed leverage
sampling_fixed_rho <- as.numeric(getOption("heston_bns_sampling_fixed_rho", -0.7))
sampling_minutes_grid <- sort(unique(as.integer(getOption(
  "heston_bns_sampling_minutes_grid",
  c(1L, 2L, 5L, 10L, 15L, 30L, 60L)
))))
sampling_base_steps_per_day <- as.integer(getOption("heston_bns_sampling_base_steps_per_day", 1440L))
n_paths_sampling_grid <- as.integer(getOption("heston_bns_n_paths_sampling_grid", 80L))
n_days_sampling_grid <- as.integer(getOption("heston_bns_n_days_sampling_grid", n_days))

if (any(!is.finite(sampling_minutes_grid)) || any(sampling_minutes_grid < 1L)) {
  stop("sampling_minutes_grid must be positive integers.")
}

scenario_tbl
```

# 4) Heston No-Jump Simulator (Full-Truncation Euler)

```{r simulator}
simulate_heston_no_jump <- function(
  n_days,
  steps_per_day,
  kappa,
  theta,
  xi,
  v0,
  rho,
  start_time = as.POSIXct("2025-01-01 00:00:00", tz = "UTC")
) {
  n_steps <- n_days * steps_per_day
  dt <- 1 / (252 * steps_per_day)
  if ((24L * 60L * 60L) %% steps_per_day != 0L) {
    stop("steps_per_day must divide 86400 exactly.")
  }
  step_seconds <- as.integer(24 * 60 * 60 / steps_per_day)

  z_w <- rnorm(n_steps)
  z_i <- rnorm(n_steps)
  z_b <- rho * z_w + sqrt(1 - rho^2) * z_i

  x <- numeric(n_steps + 1L)
  v <- numeric(n_steps + 1L)
  v[1] <- v0

  for (i in seq_len(n_steps)) {
    v_pos <- max(v[i], 0)
    dW <- sqrt(dt) * z_w[i]
    dB <- sqrt(dt) * z_b[i]

    x[i + 1L] <- x[i] + sqrt(v_pos) * dW
    v_next <- v[i] + kappa * (theta - v_pos) * dt + xi * sqrt(v_pos) * dB
    v[i + 1L] <- max(v_next, 0)
  }

  dt_index <- start_time + seq.int(0L, n_steps) * step_seconds
  out <- data.table(
    DT = as.POSIXct(dt_index, origin = "1970-01-01", tz = "UTC"),
    PRICE = exp(x),
    V = v
  )

  out[, DATE := as.Date(DT, tz = "UTC")]
  out <- out[, .(DT, PRICE, V, DATE)]
  out
}
```

# 5) BNS Helpers

```{r bns-helpers}
extract_daily_bns_summary <- function(obj, confidence_level, tz = "UTC") {
  sig_level <- 1 - confidence_level

  if (is.list(obj) && !xts::is.xts(obj) && length(obj) > 0L && is.list(obj[[1]])) {
    out <- rbindlist(lapply(seq_along(obj), function(i) {
      x <- obj[[i]]
      crit <- as.numeric(x$critical.value)
      dname <- if (!is.null(names(obj))) names(obj)[i] else NA_character_
      data.table(
        DATE = suppressWarnings(as.Date(dname)),
        ztest = as.numeric(x$ztest)[1],
        critical_lower = if (length(crit) >= 1L) crit[1] else NA_real_,
        critical_upper = if (length(crit) >= 2L) crit[2] else NA_real_,
        pvalue = as.numeric(x$pvalue)[1]
      )
    }), fill = TRUE)
  } else if (xts::is.xts(obj)) {
    raw_tbl <- data.table(
      DATE = as.Date(zoo::index(obj), tz = tz),
      as.data.frame(zoo::coredata(obj), check.names = FALSE)
    )
    p_col <- grep("^p.?value$|p-value|pvalue", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]
    z_col <- grep("^z.?test$|^ztest$", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]
    lo_col <- grep("^lower$|critical.*lower", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]
    up_col <- grep("^upper$|critical.*upper", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]

    out <- raw_tbl[, .(
      DATE,
      ztest = as.numeric(get(z_col)),
      critical_lower = as.numeric(get(lo_col)),
      critical_upper = as.numeric(get(up_col)),
      pvalue = as.numeric(get(p_col))
    )]
  } else {
    stop("Unsupported BNSjumpTest output class: ", paste(class(obj), collapse = ", "))
  }

  out <- out[!is.na(DATE)]
  out[, `:=`(
    confidence_level = confidence_level,
    nominal_alpha = 1 - confidence_level,
    reject_null = !is.na(pvalue) & (pvalue < (1 - confidence_level))
  )]
  setorder(out, DATE)
  out
}

run_daily_bns_grid <- function(price_dt, confidence_grid) {
  valid_dt <- copy(price_dt)
  valid_dt[, DATE := as.Date(DT, tz = "UTC")]
  valid_dates <- valid_dt[, .N, by = DATE][N > 1L, DATE]
  valid_dt <- valid_dt[DATE %in% valid_dates, .(DT, PRICE)]

  rbindlist(lapply(confidence_grid, function(conf_lvl) {
    alpha_q <- 1 - (1 - conf_lvl) / 2
    bns_obj <- BNSjumpTest(
      rData = valid_dt,
      type = "ratio",
      max = TRUE,
      makeReturns = TRUE,
      alpha = alpha_q
    )
    out <- extract_daily_bns_summary(bns_obj, confidence_level = conf_lvl, tz = "UTC")
    out[, alpha_quantile := alpha_q]
    out
  }), fill = TRUE)
}

resample_to_minutes <- function(price_dt, sample_minutes, tz = "UTC") {
  if (sample_minutes <= 1L) {
    out <- copy(price_dt)
    setorder(out, DT)
    return(out[, .(DT, PRICE)])
  }

  out <- aggregatePrice(
    pData = copy(price_dt),
    alignBy = "minutes",
    alignPeriod = sample_minutes,
    marketOpen = "00:00:00",
    marketClose = "23:59:59",
    tz = tz
  )
  out <- as.data.table(out)[, .(DT, PRICE)]
  out <- out[is.finite(PRICE) & !is.na(DT)]
  setorder(out, DT)
  out
}

calc_binom_ci <- function(x, n, conf.level = 0.95) {
  ci <- stats::binom.test(x = x, n = n, conf.level = conf.level)$conf.int
  c(lower = ci[1], upper = ci[2])
}

summarize_rejection_rates <- function(result_dt, group_cols) {
  out <- result_dt[, .(
    n_tests = .N,
    reject_count = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE),
    mean_pvalue = mean(pvalue, na.rm = TRUE)
  ), by = group_cols]

  ci_mat <- t(mapply(calc_binom_ci, out$reject_count, out$n_tests))
  out[, ci_lower := as.numeric(ci_mat[, 1L])]
  out[, ci_upper := as.numeric(ci_mat[, 2L])]
  out[, over_reject_point := rejection_rate > nominal_alpha]
  out[, over_reject_ci := ci_lower > nominal_alpha]
  out[, size_ratio := rejection_rate / nominal_alpha]
  out
}
```

# 6) Monte Carlo Experiment

```{r run-mc}
run_one_path <- function(path_id, scenario_name, rho, n_days, steps_per_day, par_list, conf_grid) {
  sim <- simulate_heston_no_jump(
    n_days = n_days,
    steps_per_day = steps_per_day,
    kappa = par_list$kappa,
    theta = par_list$theta,
    xi = par_list$xi,
    v0 = par_list$v0,
    rho = rho
  )

  bns_daily <- run_daily_bns_grid(sim[, .(DT, PRICE)], confidence_grid = conf_grid)
  bns_daily[, `:=`(
    scenario = scenario_name,
    rho = rho,
    path_id = path_id
  )]
  bns_daily
}

all_results <- rbindlist(lapply(seq_len(n_paths), function(pid) {
  rbindlist(lapply(seq_len(nrow(scenario_tbl)), function(j) {
    run_one_path(
      path_id = pid,
      scenario_name = scenario_tbl$scenario[j],
      rho = scenario_tbl$rho[j],
      n_days = n_days,
      steps_per_day = steps_per_day,
      par_list = heston_par,
      conf_grid = confidence_grid
    )
  }), fill = TRUE)
}), fill = TRUE)

cat("Total daily tests:", nrow(all_results), "\n")
head(all_results)
```

# 7) Rejection Rates vs Nominal Size

```{r rejection-summary}
summary_tbl <- summarize_rejection_rates(
  all_results,
  group_cols = c("scenario", "rho", "confidence_level", "nominal_alpha")
)

setorder(summary_tbl, confidence_level, scenario)
summary_tbl[]
```

# 8) Leverage vs No-Leverage Comparison

```{r comparison}
comp <- merge(
  summary_tbl[scenario == "No leverage", .(
    confidence_level,
    nominal_alpha,
    rejection_rate_no_lev = rejection_rate,
    ci_lower_no_lev = ci_lower,
    ci_upper_no_lev = ci_upper
  )],
  summary_tbl[scenario == "Leverage", .(
    confidence_level,
    nominal_alpha,
    rejection_rate_lev = rejection_rate,
    ci_lower_lev = ci_lower,
    ci_upper_lev = ci_upper
  )],
  by = c("confidence_level", "nominal_alpha"),
  all = TRUE
)

comp[, `:=`(
  diff_lev_minus_no_lev = rejection_rate_lev - rejection_rate_no_lev,
  lev_size_ratio = rejection_rate_lev / nominal_alpha,
  no_lev_size_ratio = rejection_rate_no_lev / nominal_alpha
)]

setorder(comp, confidence_level)
comp[]
```

# 9) Conclusion under the No-Jump Null

```{r conclusion}
conclusion_tbl <- comp[, .(
  confidence_level,
  nominal_alpha,
  rejection_rate_no_lev,
  rejection_rate_lev,
  diff_lev_minus_no_lev
)]

print(conclusion_tbl)

if (all(comp$diff_lev_minus_no_lev > 0, na.rm = TRUE)) {
  cat(
    "\nConclusion: in this simulation, leverage (negative rho) consistently increases",
    "false jump rejections versus rho = 0 under the no-jump null.\n"
  )
} else {
  cat(
    "\nConclusion: leverage does not uniformly increase false rejections",
    "for all confidence levels in this run.\n"
  )
}

if (any(summary_tbl$scenario == "Leverage" & summary_tbl$over_reject_point, na.rm = TRUE)) {
  cat(
    "At least one leverage configuration shows point-estimate over-rejection",
    "(empirical size above nominal alpha).\n"
  )
}
```

# 10) Effect of Leverage Level (How Negative rho Is)

```{r leverage-level-run}
leverage_level_results <- rbindlist(lapply(seq_len(n_paths_leverage_grid), function(pid) {
  rbindlist(lapply(leverage_rho_grid, function(rho_i) {
    out <- run_one_path(
      path_id = pid,
      scenario_name = sprintf("rho=%.2f", rho_i),
      rho = rho_i,
      n_days = n_days_leverage_grid,
      steps_per_day = steps_per_day,
      par_list = heston_par,
      conf_grid = confidence_grid
    )
    out[, rho_level := rho_i]
    out
  }), fill = TRUE)
}), fill = TRUE)

cat("Leverage-level daily tests:", nrow(leverage_level_results), "\n")
```

```{r leverage-level-summary}
leverage_level_summary <- summarize_rejection_rates(
  leverage_level_results,
  group_cols = c("rho_level", "confidence_level", "nominal_alpha")
)

setorder(leverage_level_summary, confidence_level, -rho_level)
leverage_level_summary[]
```

```{r leverage-level-trend}
leverage_trend <- leverage_level_summary[, {
  ord <- order(abs(rho_level))
  rr_ord <- rejection_rate[ord]
  rho_ord <- abs(rho_level[ord])
  data.table(
    corr_abs_rho_vs_rejection = suppressWarnings(cor(rho_ord, rr_ord)),
    slope_abs_rho = as.numeric(coef(lm(rr_ord ~ rho_ord))[2]),
    monotone_non_decreasing = all(diff(rr_ord) >= -1e-12)
  )
}, by = .(confidence_level, nominal_alpha)]

setorder(leverage_trend, confidence_level)
leverage_trend[]
```

# 11) Effect of Sampling Interval (Fixed Leverage)

```{r sampling-grid-run}
sampling_grid_results <- rbindlist(lapply(seq_len(n_paths_sampling_grid), function(pid) {
  sim_path <- simulate_heston_no_jump(
    n_days = n_days_sampling_grid,
    steps_per_day = sampling_base_steps_per_day,
    kappa = heston_par$kappa,
    theta = heston_par$theta,
    xi = heston_par$xi,
    v0 = heston_par$v0,
    rho = sampling_fixed_rho
  )

  rbindlist(lapply(sampling_minutes_grid, function(m) {
    sampled <- resample_to_minutes(sim_path[, .(DT, PRICE)], sample_minutes = m, tz = "UTC")
    out <- run_daily_bns_grid(sampled, confidence_grid = confidence_grid)
    out[, `:=`(
      path_id = pid,
      rho = sampling_fixed_rho,
      sample_minutes = m
    )]
    out
  }), fill = TRUE)
}), fill = TRUE)

cat("Sampling-grid daily tests:", nrow(sampling_grid_results), "\n")
```

```{r sampling-grid-summary}
sampling_grid_summary <- summarize_rejection_rates(
  sampling_grid_results,
  group_cols = c("sample_minutes", "rho", "confidence_level", "nominal_alpha")
)

setorder(sampling_grid_summary, confidence_level, sample_minutes)
sampling_grid_summary[]
```

```{r sampling-grid-trend}
sampling_trend <- sampling_grid_summary[, {
  ord <- order(sample_minutes)
  rr_ord <- rejection_rate[ord]
  m_ord <- sample_minutes[ord]
  data.table(
    corr_minutes_vs_rejection = suppressWarnings(cor(m_ord, rr_ord)),
    slope_minutes = as.numeric(coef(lm(rr_ord ~ m_ord))[2]),
    monotone_non_increasing_when_finer = all(diff(rev(rr_ord)) <= 1e-12)
  )
}, by = .(confidence_level, nominal_alpha)]

setorder(sampling_trend, confidence_level)
sampling_trend[]
```

# 12) Focused Conclusion for New Questions

```{r focused-conclusion}
cat("Leverage-level trend (more negative rho means larger |rho|):\n")
print(leverage_trend)

cat("\nSampling-interval trend (fixed rho =", sampling_fixed_rho, "):\n")
print(sampling_trend)

cat(
  "\nInterpretation guide:\n",
  "- Positive slope_abs_rho: stronger leverage (more negative rho) increases false rejections.\n",
  "- Negative slope_minutes: finer sampling (smaller minute interval) increases false rejections.\n",
  sep = ""
)
```
