---
title: "BTC Long-Horizon Daily Jump Tests (Binance 1m)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# 1) Install and Load Packages

```{r packages}
required_pkgs <- c("highfrequency", "data.table", "xts", "zoo")
missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, logical(1), quietly = TRUE)]

if (length(missing_pkgs) > 0L) {
  stop(
    "Missing required package(s): ",
    paste(missing_pkgs, collapse = ", "),
    ". Install them before rendering this Rmd."
  )
}

library(highfrequency)
library(data.table)
library(xts)
library(zoo)
```

# 2) Load BTC Binance 1m Data

```{r load-data}
input_path <- "data/okx_1m/BTCUSDT_1m_2025-01-01_to_2026-05-31.csv.gz"
if (!file.exists(input_path)) {
  stop("Missing input file: ", input_path)
}

if (requireNamespace("R.utils", quietly = TRUE)) {
  raw_btc <- fread(
    input_path,
    select = c("open_time", "open_time_ms", "close")
  )
} else {
  raw_btc <- as.data.table(
    read.csv(
      gzfile(input_path),
      stringsAsFactors = FALSE
    )
  )[, .(open_time, open_time_ms, close)]
}

raw_btc[, DT := suppressWarnings(as.POSIXct(open_time, tz = "UTC"))]
if (all(is.na(raw_btc$DT))) {
  raw_btc[, DT := as.POSIXct(open_time_ms / 1000, origin = "1970-01-01", tz = "UTC")]
}

raw_btc[, PRICE := as.numeric(close)]
pData <- raw_btc[
  is.finite(PRICE) & PRICE > 0 & !is.na(DT),
  .(DT, PRICE)
]
setorder(pData, DT)

cat("Input file:", input_path, "\n")
cat("Rows after cleanup:", nrow(pData), "\n")
cat("Date range:", format(min(pData$DT), "%Y-%m-%d %H:%M:%S"), "to", format(max(pData$DT), "%Y-%m-%d %H:%M:%S"), "UTC\n")
head(pData)
```

```{r}
pData
```

# 3) Parameters and Helper Functions

```{r helpers}
sample_freq_grid <- c(1L, 2L, 5L, 15L, 30L, 60L, 120L)
confidence_grid <- c(0.90, 0.95, 0.99)
default_confidence <- 0.95
acf_lag_max <- 5L
plot_freqs <- c(1L, 5L, 30L, 120L)

resample_prices <- function(pdata, freq_mins, tz = "UTC") {
  out <- copy(pdata)
  if (freq_mins > 1L) {
    out <- aggregatePrice(
      out,
      alignBy = "minutes",
      alignPeriod = freq_mins,
      marketOpen = "00:00:00",
      marketClose = "23:59:59",
      tz = tz
    )
  }

  out <- as.data.table(out)
  out <- out[is.finite(PRICE) & PRICE > 0 & !is.na(DT)]
  setorder(out, DT)

  out[, DATE := as.Date(DT, tz = tz)]
  valid_dates <- out[, .N, by = DATE][N > 1L, DATE]
  out <- out[DATE %in% valid_dates, .(DT, PRICE)]
  setorder(out, DT)
  out
}

compute_return_acf_summary <- function(sampled_pdata, lag_max = 5L) {
  tmp <- copy(sampled_pdata)
  tmp[, ret := log(PRICE) - shift(log(PRICE), type = "lag")]
  ret <- tmp[is.finite(ret), ret]

  if (length(ret) < (lag_max + 1L)) {
    return(data.table(
      n_returns = length(ret),
      lag1_acf = NA_real_,
      lag1_to_lag5_acf_sum = NA_real_,
      lag1_to_lag5_acf_abs_sum = NA_real_
    ))
  }

  acf_obj <- stats::acf(ret, plot = FALSE, lag.max = lag_max, na.action = na.pass)
  lag_vals <- as.numeric(acf_obj$acf[2:(lag_max + 1L)])

  data.table(
    n_returns = length(ret),
    lag1_acf = lag_vals[1],
    lag1_to_lag5_acf_sum = sum(lag_vals, na.rm = TRUE),
    lag1_to_lag5_acf_abs_sum = sum(abs(lag_vals), na.rm = TRUE)
  )
}

to_numeric_xts <- function(x) {
  as.numeric(zoo::coredata(x))
}

daywise_noise_robust <- function(sampled_pdata, estimator = c("rTSCov", "rRTSCov"), tz = "UTC") {
  estimator <- match.arg(estimator)
  px_xts <- xts(sampled_pdata$PRICE, order.by = sampled_pdata$DT, tzone = tz)
  day_index <- as.Date(index(px_xts), tz = tz)
  split_rows <- split(seq_along(day_index), day_index)

  out <- rbindlist(lapply(names(split_rows), function(d) {
    rows <- split_rows[[d]]
    x_day <- px_xts[rows]
    n_obs <- NROW(x_day)

    if (n_obs < 20L) {
      return(data.table(DATE = as.Date(d), value = NA_real_, n_obs = n_obs))
    }

    K_use <- max(2L, floor((n_obs - 1L) / 10L))
    K_use <- min(K_use, 50L)
    J_use <- 1L

    est_val <- tryCatch(
      {
        if (estimator == "rTSCov") {
          as.numeric(rTSCov(x_day, K = K_use, J = J_use))
        } else {
          as.numeric(rRTSCov(x_day, K = K_use, J = J_use))
        }
      },
      error = function(e) NA_real_
    )

    data.table(DATE = as.Date(d), value = est_val, n_obs = n_obs)
  }), fill = TRUE)

  setorder(out, DATE)
  out
}

extract_daily_bns_summary <- function(obj, confidence_level, tz = "UTC") {
  sig_level <- 1 - confidence_level

  if (is.list(obj) && !xts::is.xts(obj) && length(obj) > 0 && is.list(obj[[1]])) {
    out <- rbindlist(lapply(seq_along(obj), function(i) {
      x <- obj[[i]]
      crit <- as.numeric(x$critical.value)
      dname <- if (!is.null(names(obj))) names(obj)[i] else NA_character_
      data.table(
        DATE = suppressWarnings(as.Date(dname)),
        ztest = as.numeric(x$ztest)[1],
        critical_lower = if (length(crit) >= 1) crit[1] else NA_real_,
        critical_upper = if (length(crit) >= 2) crit[2] else NA_real_,
        pvalue = as.numeric(x$pvalue)[1]
      )
    }), fill = TRUE)
  } else if (xts::is.xts(obj)) {
    raw_tbl <- data.table(
      DATE = as.Date(zoo::index(obj), tz = tz),
      as.data.frame(zoo::coredata(obj), check.names = FALSE)
    )
    p_col <- grep("^p.?value$|p-value|pvalue", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]
    z_col <- grep("^z.?test$|^ztest$", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]
    lo_col <- grep("^lower$|critical.*lower", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]
    up_col <- grep("^upper$|critical.*upper", names(raw_tbl), ignore.case = TRUE, value = TRUE)[1]

    out <- raw_tbl[, .(
      DATE,
      ztest = as.numeric(get(z_col)),
      critical_lower = as.numeric(get(lo_col)),
      critical_upper = as.numeric(get(up_col)),
      pvalue = as.numeric(get(p_col))
    )]
  } else {
    stop("Unsupported BNSjumpTest output class: ", paste(class(obj), collapse = ", "))
  }

  out[, reject_null := !is.na(pvalue) & (pvalue < sig_level)]
  out[, confidence_level := confidence_level]
  setorder(out, DATE)
  out
}

run_daily_bns <- function(sampled_pdata, confidence_level) {
  alpha_quantile <- 1 - (1 - confidence_level) / 2

  # Default structure requested by user:
  # daily_bns <- BNSjumpTest(
  #   rData = pData_daily_test,
  #   type = "ratio",
  #   max = TRUE,
  #   makeReturns = TRUE
  # )
  bns_obj <- BNSjumpTest(
    rData = sampled_pdata,
    type = "ratio",
    max = TRUE,
    makeReturns = TRUE,
    alpha = alpha_quantile
  )

  extract_daily_bns_summary(
    bns_obj,
    confidence_level = confidence_level,
    tz = "UTC"
  )
}
```

# 4) Run Long-Horizon Frequency/Confidence Grid

```{r run-grid}
grid_results <- lapply(sample_freq_grid, function(freq_mins) {
  sampled <- resample_prices(pData, freq_mins = freq_mins, tz = "UTC")
  acf_summary <- compute_return_acf_summary(sampled, lag_max = acf_lag_max)
  acf_summary[, sample_minutes := freq_mins]

  daily_details <- rbindlist(lapply(confidence_grid, function(conf_lvl) {
    daily <- run_daily_bns(sampled, confidence_level = conf_lvl)
    daily[, `:=`(
      sample_minutes = freq_mins,
      alpha_quantile = 1 - (1 - conf_lvl) / 2
    )]
    daily
  }), fill = TRUE)

  jump_summary <- daily_details[, .(
    n_test_days = .N,
    reject_days = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE),
    mean_ratio_stat = mean(ztest, na.rm = TRUE),
    mean_critical_lower = mean(critical_lower, na.rm = TRUE),
    mean_pvalue = mean(pvalue, na.rm = TRUE)
  ), by = .(sample_minutes, confidence_level, alpha_quantile)]

  list(
    acf_summary = acf_summary,
    daily_details = daily_details,
    jump_summary = jump_summary
  )
})

acf_summary_all <- rbindlist(lapply(grid_results, `[[`, "acf_summary"), fill = TRUE)
daily_details_all <- rbindlist(lapply(grid_results, `[[`, "daily_details"), fill = TRUE)
jump_summary_all <- rbindlist(lapply(grid_results, `[[`, "jump_summary"), fill = TRUE)

setorder(acf_summary_all, sample_minutes)
setorder(jump_summary_all, sample_minutes, confidence_level)
setorder(daily_details_all, sample_minutes, confidence_level, DATE)
```

# 5) Autocorrelation Diagnostics by Sampling Frequency

```{r acf-table}
acf_summary_all
```

```{r acf-plot}
barplot(
  height = acf_summary_all$lag1_to_lag5_acf_sum,
  names.arg = acf_summary_all$sample_minutes,
  xlab = "Sampling Frequency (minutes)",
  ylab = "Sum of ACF (lags 1 to 5)",
  main = "BTC Return Serial Correlation by Sampling Frequency",
  col = "steelblue"
)
abline(h = 0, lty = 2, col = "gray40")
```

```{r microstructure-summary-table}
microstructure_summary_all <- rbindlist(lapply(sample_freq_grid, function(freq_mins) {
  sampled <- resample_prices(pData, freq_mins = freq_mins, tz = "UTC")

  sampled[, ret := log(PRICE) - shift(log(PRICE), type = "lag")]
  ret_vec <- sampled[is.finite(ret), ret]

  lag1_acf <- if (length(ret_vec) > 1L) {
    cor(ret_vec[-1], ret_vec[-length(ret_vec)])
  } else {
    NA_real_
  }

  ljung_p <- if (length(ret_vec) > 20L) {
    Box.test(ret_vec, lag = 10L, type = "Ljung-Box")$p.value
  } else {
    NA_real_
  }

  px_xts <- xts(sampled$PRICE, order.by = sampled$DT, tzone = "UTC")
  rv_xts <- rCov(px_xts, makeReturns = TRUE)
  bp_xts <- rBPCov(px_xts, makeReturns = TRUE)
  rk_xts <- suppressWarnings(rKernelCov(px_xts, makeReturns = TRUE))

  rv_vec <- to_numeric_xts(rv_xts)
  bp_vec <- to_numeric_xts(bp_xts)
  rk_vec <- to_numeric_xts(rk_xts)
  rk_vec[rk_vec < 0] <- NA_real_

  ts_tbl <- daywise_noise_robust(sampled, estimator = "rTSCov", tz = "UTC")
  rts_tbl <- daywise_noise_robust(sampled, estimator = "rRTSCov", tz = "UTC")

  data.table(
    sample_minutes = freq_mins,
    n_intraday_returns = length(ret_vec),
    lag1_return_acf = lag1_acf,
    ljung_box_pvalue_lag10 = ljung_p,
    mean_daily_rv = mean(rv_vec, na.rm = TRUE),
    mean_daily_bp = mean(bp_vec, na.rm = TRUE),
    mean_daily_rk = mean(rk_vec, na.rm = TRUE),
    mean_daily_ts = mean(ts_tbl$value, na.rm = TRUE),
    mean_daily_rts = mean(rts_tbl$value, na.rm = TRUE),
    rv_minus_rk = mean(rv_vec - rk_vec, na.rm = TRUE),
    rv_minus_rts = mean(rv_vec - rts_tbl$value, na.rm = TRUE)
  )
}), fill = TRUE)

setorder(microstructure_summary_all, sample_minutes)
microstructure_summary_all
```

```{r microstructure-signature-plot}
plot(
  microstructure_summary_all$sample_minutes,
  microstructure_summary_all$mean_daily_rv,
  type = "b",
  pch = 19,
  col = "firebrick",
  xlab = "Sampling Frequency (minutes)",
  ylab = "Mean Daily RV",
  main = "Volatility Signature Plot"
)
```

```{r microstructure-lag1-plot}
plot(
  microstructure_summary_all$sample_minutes,
  microstructure_summary_all$lag1_return_acf,
  type = "b",
  pch = 19,
  col = "darkgreen",
  xlab = "Sampling Frequency (minutes)",
  ylab = "Lag-1 Return ACF",
  main = "Lag-1 Return Autocorrelation by Sampling Frequency"
)
abline(h = 0, lty = 2, col = "gray40")
```

```{r microstructure-noise-gap-plot}
plot(
  microstructure_summary_all$sample_minutes,
  microstructure_summary_all$rv_minus_rk,
  type = "b",
  pch = 19,
  col = "navy",
  xlab = "Sampling Frequency (minutes)",
  ylab = "Mean(RV - RK)",
  main = "Noise Gap by Sampling Frequency"
)
abline(h = 0, lty = 2, col = "gray40")
```

# 5.5) Leverage Effects

```{r leverage-effects-summary}
leverage_summary_all <- rbindlist(lapply(sample_freq_grid, function(freq_mins) {
  sampled <- resample_prices(pData, freq_mins = freq_mins, tz = "UTC")
  sampled[, DATE := as.Date(DT, tz = "UTC")]

  daily_close <- sampled[, .(close_price = PRICE[.N]), by = DATE][order(DATE)]
  daily_close[, ret_t := log(close_price) - shift(log(close_price), type = "lag")]

  px_xts <- xts(sampled$PRICE, order.by = sampled$DT, tzone = "UTC")
  rv_xts <- rCov(px_xts, makeReturns = TRUE)
  vol_tbl <- data.table(
    DATE = as.Date(zoo::index(rv_xts), tz = "UTC"),
    rv = to_numeric_xts(rv_xts)
  )
  vol_tbl[, vol_t := sqrt(pmax(rv, 0))]
  vol_tbl[, dvol_t_to_t1 := shift(vol_t, type = "lead") - vol_t]

  lev_tbl <- merge(
    daily_close[, .(DATE, ret_t)],
    vol_tbl[, .(DATE, vol_t, dvol_t_to_t1)],
    by = "DATE",
    all = FALSE
  )

  valid <- lev_tbl[is.finite(ret_t) & is.finite(dvol_t_to_t1)]
  n_pairs <- nrow(valid)

  leverage_corr <- if (n_pairs >= 3L) {
    cor(valid$ret_t, valid$dvol_t_to_t1)
  } else {
    NA_real_
  }

  leverage_pvalue <- if (n_pairs >= 4L) {
    tryCatch(
      cor.test(valid$ret_t, valid$dvol_t_to_t1)$p.value,
      error = function(e) NA_real_
    )
  } else {
    NA_real_
  }

  data.table(
    sample_minutes = freq_mins,
    n_pairs = n_pairs,
    leverage_corr = leverage_corr,
    leverage_pvalue = leverage_pvalue,
    mean_ret_t = mean(valid$ret_t, na.rm = TRUE),
    mean_dvol = mean(valid$dvol_t_to_t1, na.rm = TRUE)
  )
}), fill = TRUE)

setorder(leverage_summary_all, sample_minutes)
leverage_summary_all
```

```{r leverage-effects-plot}
plot(
  leverage_summary_all$sample_minutes,
  leverage_summary_all$leverage_corr,
  type = "b",
  pch = 19,
  col = "brown3",
  xlab = "Sampling Frequency (minutes)",
  ylab = "Corr(r_t, vol_{t+1} - vol_t)",
  main = "Leverage Effect by Sampling Frequency"
)
abline(h = 0, lty = 2, col = "gray40")
```

# 6) Rejection Rate over Confidence and Sampling Frequency


```{r rejection-summary}
jump_summary_all
```

```{r}
daily_details_all[confidence_level == 0.95 & sample_minutes == 30 & reject_null == TRUE]
```

```{r rejection-rate-by-frequency-plot}
plot_confidence_levels <- sort(unique(jump_summary_all$confidence_level))
old_par <- par(no.readonly = TRUE)
on.exit(par(old_par))

# par(mfrow = c(length(plot_alpha_levels), 1), mar = c(3.5, 4, 2.5, 1))

for (c_q in plot_confidence_levels) {
  panel <- jump_summary_all[confidence_level == c_q & sample_minutes >= 5][order(sample_minutes)]
  y_max <- suppressWarnings(max(panel$rejection_rate, na.rm = TRUE))
  if (!is.finite(y_max)) {
    y_max <- 0.1
  }
  y_lim <- c(0, max(0.05, y_max * 1.10))

  plot(
    panel$sample_minutes,
    panel$rejection_rate,
    type = "b",
    pch = 19,
    col = "steelblue4",
    ylim = y_lim,
    xlab = "Sampling Frequency (minutes)",
    ylab = "Rejection Rate",
    main = paste0(
      "Rejection Rate by Sampling Frequency (confidence level = ",
      sprintf("%.3f", c_q),
      ")"
    )
  )
  abline(h = 0, lty = 2, col = "gray50")
}
```

# 7) Figure Similar to Reference (Ratio Statistic vs Critical Value)

```{r ratio-critical-plot}

```


```{r ratio-critical-plot}
plot_confidence <- 0.95
plot_freqs <- c(1L, 5L, 30L, 120L)

plot_data <- daily_details_all[
  sample_minutes %in% plot_freqs &
    abs(confidence_level - plot_confidence) < 1e-10
]

plot_data <- plot_data[order(sample_minutes, DATE)]

old_par <- par(no.readonly = TRUE)
on.exit(par(old_par))

# par(mfrow = c(2, 2), mar = c(3, 3, 2.5, 1), oma = c(1.2, 1.2, 0, 0))

for (i in seq_along(plot_freqs)) {
  freq <- plot_freqs[i]
  panel <- plot_data[sample_minutes == freq]

  if (nrow(panel) == 0) {
    plot.new()
    text(0.5, 0.5, paste0("No data at ", freq, "m"))
    next
  }

  x <- seq_len(nrow(panel))
  y_stat <- panel$ztest
  y_stat_plot <- pmax(y_stat, 0)
  y_cv <- panel$critical_upper

  y_cv_max <- suppressWarnings(max(y_cv, na.rm = TRUE))
  if (!is.finite(y_cv_max)) {
    y_cv_max <- 1
  }
  y_upper <- y_cv_max * 3

  plot(
    x, y_stat_plot,
    type = "n",
    ylim = c(0, y_upper),
    xlab = "",
    ylab = "",
    main = paste0("Sampling = ", freq, " min (conf = ", plot_confidence, ")"),
    bty = "l"
  )

  # segments(
  #   x0 = x,
  #   y0 = 0,
  #   x1 = x,
  #   y1 = y_stat_plot,
  #   col = "gray30",
  #   lwd = 0.8
  # )
  points(x, y_stat_plot, pch = 3, col = "gray20", cex = 0.7)
  lines(x, y_cv, col = "gray20", lwd = 1.2)
  abline(h = 0, col = "gray20", lwd = 1.5)

  if (i %% 2 == 1L) {
    mtext("Statistic", side = 2, line = 2)
  }
  if (i > 2L) {
    mtext("Days", side = 1, line = 2)
  }
  # if (i == 1L) {
  #   legend(
  #     "topright",
  #     legend = c("Z-test statistic (positive part)", "Positive critical value"),
  #     pch = c(3, NA),
  #     lty = c(NA, 1),
  #     col = c("gray20", "gray20"),
  #     bty = "n",
  #     cex = 0.9
  #   )
  # }
}
```

# 8) Final Combined Table (Autocorrelation + Rejection Rate)

```{r final-table}
final_table <- merge(
  jump_summary_all,
  acf_summary_all,
  by = "sample_minutes",
  all.x = TRUE
)

setorder(final_table, sample_minutes, confidence_level)

final_table[
  ,
  .(
    sample_minutes,
    confidence_level,
    alpha_quantile,
    n_returns,
    lag1_acf,
    lag1_to_lag5_acf_sum,
    n_test_days,
    reject_days,
    rejection_rate
  )
]
```

# 9) Period-Based Rejection and Largest-Ztest Day (30m, 95%)

```{r period-controls}
# Fixed context requested:
fixed_confidence_level <- 0.95
fixed_sample_minutes <- 30L

# User control: choose "month", "quarter", or "all".
period_unit <- "month"

# Optional plotting controls:
# - plot_sample_minutes controls only section-9 visualization frequency.
#   (Temporarily set to 1-minute as requested.)
plot_sample_minutes <- 1L
#
# - If NULL, uses the first available period.
# - If visualized_day is NULL, uses the max-ztest day in selected period.
period_to_plot <- NULL
visualized_day <- NULL

daily_tests <- daily_details_all[
  abs(confidence_level - fixed_confidence_level) < 1e-10 &
    sample_minutes == fixed_sample_minutes
]

if (nrow(daily_tests) == 0) {
  stop(
    "No daily tests found for confidence_level = ",
    fixed_confidence_level,
    " and sample_minutes = ",
    fixed_sample_minutes
  )
}

add_period_columns <- function(dt, unit = c("month", "quarter", "all")) {
  unit <- match.arg(unit)
  out <- copy(dt)

  if (unit == "month") {
    out[, period_id := format(DATE, "%Y-%m")]
    out[, period_start := as.Date(format(DATE, "%Y-%m-01"))]
  } else if (unit == "quarter") {
    out[, period_year := as.integer(format(DATE, "%Y"))]
    out[, period_quarter := (as.integer(format(DATE, "%m")) - 1L) %/% 3L + 1L]
    out[, period_id := sprintf("%d-Q%d", period_year, period_quarter)]
    out[, period_start := as.Date(
      sprintf("%d-%02d-01", period_year, (period_quarter - 1L) * 3L + 1L)
    )]
  } else {
    out[, period_id := "ALL"]
    out[, period_start := min(DATE, na.rm = TRUE)]
  }

  out[, period_end := max(DATE), by = .(period_id, period_start)]
  setorder(out, period_start, DATE)
  out
}

periodized_daily_tests <- add_period_columns(daily_tests, unit = period_unit)

cat("Fixed confidence level:", fixed_confidence_level, "\n")
cat("Fixed sample frequency:", fixed_sample_minutes, "minutes\n")
cat("Period grouping:", period_unit, "\n")
cat("Plot sampling frequency:", plot_sample_minutes, "minutes\n")
```

```{r rejection-rate-by-period-table}
rejection_rate_by_period <- periodized_daily_tests[
  ,
  .(
    period_start = min(DATE),
    period_end = max(DATE),
    n_test_days = .N,
    reject_days = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE)
  ),
  by = period_id
]

setorder(rejection_rate_by_period, period_start)
rejection_rate_by_period[]
```

```{r rejection-vs-daily-realized-volatility}
# Daily realized variance/volatility at the same test sampling frequency.
daily_rv_sampled <- resample_prices(
  pData,
  freq_mins = fixed_sample_minutes,
  tz = "UTC"
)

daily_rv_xts <- rCov(
  xts(daily_rv_sampled$PRICE, order.by = daily_rv_sampled$DT, tzone = "UTC"),
  makeReturns = TRUE
)

daily_rv_tbl <- data.table(
  DATE = as.Date(zoo::index(daily_rv_xts), tz = "UTC"),
  daily_rv = to_numeric_xts(daily_rv_xts)
)
daily_rv_tbl[, daily_vol := sqrt(pmax(daily_rv, 0))]

daily_reject_vs_rv <- merge(
  periodized_daily_tests[
    ,
    .(DATE, period_id, period_start, period_end, reject_null, ztest, pvalue)
  ],
  daily_rv_tbl,
  by = "DATE",
  all.x = TRUE
)

daily_reject_vs_rv <- daily_reject_vs_rv[
  is.finite(daily_rv) & is.finite(daily_vol) & !is.na(reject_null)
]
daily_reject_vs_rv[, reject_flag := as.integer(reject_null)]

daily_corr_test <- if (
  nrow(daily_reject_vs_rv) >= 3L &&
    length(unique(daily_reject_vs_rv$reject_flag)) > 1L
) {
  tryCatch(
    cor.test(daily_reject_vs_rv$reject_flag, daily_reject_vs_rv$daily_vol),
    error = function(e) NULL
  )
} else {
  NULL
}

daily_reject_rv_correlation <- data.table(
  n_days = nrow(daily_reject_vs_rv),
  reject_days = sum(daily_reject_vs_rv$reject_null, na.rm = TRUE),
  mean_daily_vol_reject = daily_reject_vs_rv[
    reject_null == TRUE,
    mean(daily_vol, na.rm = TRUE)
  ],
  mean_daily_vol_non_reject = daily_reject_vs_rv[
    reject_null == FALSE,
    mean(daily_vol, na.rm = TRUE)
  ],
  corr_reject_vs_daily_vol = if (is.null(daily_corr_test)) {
    NA_real_
  } else {
    as.numeric(daily_corr_test$estimate)
  },
  corr_pvalue = if (is.null(daily_corr_test)) {
    NA_real_
  } else {
    daily_corr_test$p.value
  }
)

daily_reject_rv_correlation[]
```

```{r rejection-vs-daily-realized-volatility-plot, fig.width=8, fig.height=4}
if (nrow(daily_reject_vs_rv) == 0L) {
  stop("No rows available for reject-vs-daily-volatility visualization.")
}

boxplot(
  daily_vol ~ reject_null,
  data = daily_reject_vs_rv,
  names = c("No reject", "Reject"),
  col = c("#9ECAE1", "#FC9272"),
  xlab = "Reject null?",
  ylab = "Daily realized volatility (sqrt(RV))",
  main = paste0(
    "Daily Rejection vs Realized Volatility (",
    fixed_sample_minutes, "-minute sampling)"
  )
)
```

```{r rejection-rate-vs-period-realized-volatility}
period_vol_table <- daily_reject_vs_rv[
  ,
  .(
    period_start = min(DATE),
    period_end = max(DATE),
    n_days = .N,
    reject_days = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE),
    period_rv_sum = sum(pmax(daily_rv, 0), na.rm = TRUE),
    period_realized_vol = sqrt(sum(pmax(daily_rv, 0), na.rm = TRUE)),
    mean_daily_vol = mean(daily_vol, na.rm = TRUE)
  ),
  by = period_id
]
setorder(period_vol_table, period_start)

period_corr_test <- if (
  nrow(period_vol_table) >= 3L &&
    length(unique(period_vol_table$rejection_rate)) > 1L &&
    length(unique(period_vol_table$period_realized_vol)) > 1L
) {
  tryCatch(
    cor.test(period_vol_table$rejection_rate, period_vol_table$period_realized_vol),
    error = function(e) NULL
  )
} else {
  NULL
}

period_reject_rv_correlation <- data.table(
  n_periods = nrow(period_vol_table),
  corr_rejection_rate_vs_period_realized_vol = if (is.null(period_corr_test)) {
    NA_real_
  } else {
    as.numeric(period_corr_test$estimate)
  },
  corr_pvalue = if (is.null(period_corr_test)) {
    NA_real_
  } else {
    period_corr_test$p.value
  }
)

period_reject_rv_correlation[]
period_vol_table[]
```

```{r highest-vs-lowest-rejection-period-price-plot, fig.width=11, fig.height=7}
period_ranked <- period_vol_table[is.finite(rejection_rate)][order(-rejection_rate, period_start)]
if (nrow(period_ranked) == 0L) {
  stop("No finite period rejection rates are available for comparison.")
}

highest_reject_period <- period_ranked[1]
lowest_reject_period <- period_ranked[.N]
selected_period_ids <- unique(c(highest_reject_period$period_id, lowest_reject_period$period_id))

period_price_sampled <- resample_prices(
  pData,
  freq_mins = plot_sample_minutes,
  tz = "UTC"
)
period_price_sampled[, DATE := as.Date(DT, tz = "UTC")]
period_price_sampled <- add_period_columns(period_price_sampled, unit = period_unit)

high_period_prices <- period_price_sampled[period_id == highest_reject_period$period_id]
low_period_prices <- period_price_sampled[period_id == lowest_reject_period$period_id]

op <- par(no.readonly = TRUE)
par(mfrow = c(2, 1), mar = c(5, 4, 4, 1))

plot(
  high_period_prices$DT,
  high_period_prices$PRICE,
  type = "l",
  col = "#D95F02",
  lwd = 1.2,
  xlab = "Time (UTC)",
  ylab = "Price",
  main = paste0(
    "Highest Rejection Period: ", highest_reject_period$period_id,
    " (rate=", sprintf("%.3f", highest_reject_period$rejection_rate), ")"
  )
)

plot(
  low_period_prices$DT,
  low_period_prices$PRICE,
  type = "l",
  col = "#1B9E77",
  lwd = 1.2,
  xlab = "Time (UTC)",
  ylab = "Price",
  main = paste0(
    "Lowest Rejection Period: ", lowest_reject_period$period_id,
    " (rate=", sprintf("%.3f", lowest_reject_period$rejection_rate), ")"
  )
)

par(op)

period_vol_table[
  period_id %in% selected_period_ids,
  .(
    period_id,
    period_start,
    period_end,
    n_days,
    reject_days,
    rejection_rate,
    period_realized_vol,
    mean_daily_vol
  )
][order(-rejection_rate, period_start)]
```

```{r microstructure-noise-diagnostics-extreme-periods}
diagnostic_period_bounds <- unique(
  period_vol_table[
    period_id %in% selected_period_ids,
    .(period_id, period_start, period_end)
  ]
)

microstructure_period_extremes <- rbindlist(lapply(seq_len(nrow(diagnostic_period_bounds)), function(i) {
  pid <- diagnostic_period_bounds$period_id[i]
  pstart <- diagnostic_period_bounds$period_start[i]
  pend <- diagnostic_period_bounds$period_end[i]

  period_start_dt <- as.POSIXct(
    paste(as.character(pstart), "00:00:00"),
    tz = "UTC"
  )
  period_end_dt <- as.POSIXct(
    paste(as.character(pend), "23:59:59"),
    tz = "UTC"
  )

  p_period <- pData[DT >= period_start_dt & DT <= period_end_dt]
  if (nrow(p_period) < 50L) {
    return(NULL)
  }

  rbindlist(lapply(sample_freq_grid, function(freq_mins) {
    sampled <- resample_prices(p_period, freq_mins = freq_mins, tz = "UTC")
    sampled[, ret := log(PRICE) - shift(log(PRICE), type = "lag")]
    ret_vec <- sampled[is.finite(ret), ret]

    if (length(ret_vec) < 20L) {
      return(data.table(
        period_id = pid,
        period_start = pstart,
        period_end = pend,
        sample_minutes = freq_mins,
        n_intraday_returns = length(ret_vec),
        lag1_return_acf = NA_real_,
        lag1_to_lag5_acf_sum = NA_real_,
        ljung_box_pvalue_lag10 = NA_real_,
        mean_daily_rv = NA_real_,
        mean_daily_bp = NA_real_,
        mean_daily_rk = NA_real_,
        mean_daily_ts = NA_real_,
        mean_daily_rts = NA_real_,
        rv_minus_rk = NA_real_,
        rv_minus_rts = NA_real_
      ))
    }

    lag1_acf <- suppressWarnings(cor(ret_vec[-1], ret_vec[-length(ret_vec)]))
    lag1_to_lag5_acf_sum <- if (length(ret_vec) >= 6L) {
      acf_obj <- stats::acf(ret_vec, plot = FALSE, lag.max = 5L, na.action = na.pass)
      sum(as.numeric(acf_obj$acf[2:6]), na.rm = TRUE)
    } else {
      NA_real_
    }
    ljung_p <- if (length(ret_vec) > 20L) {
      Box.test(ret_vec, lag = 10L, type = "Ljung-Box")$p.value
    } else {
      NA_real_
    }

    px_xts <- xts(sampled$PRICE, order.by = sampled$DT, tzone = "UTC")
    rv_xts <- rCov(px_xts, makeReturns = TRUE)
    bp_xts <- rBPCov(px_xts, makeReturns = TRUE)
    rk_xts <- suppressWarnings(rKernelCov(px_xts, makeReturns = TRUE))

    rv_vec <- to_numeric_xts(rv_xts)
    bp_vec <- to_numeric_xts(bp_xts)
    rk_vec <- to_numeric_xts(rk_xts)
    rk_vec[rk_vec < 0] <- NA_real_

    ts_tbl <- daywise_noise_robust(sampled, estimator = "rTSCov", tz = "UTC")
    rts_tbl <- daywise_noise_robust(sampled, estimator = "rRTSCov", tz = "UTC")

    data.table(
      period_id = pid,
      period_start = pstart,
      period_end = pend,
      sample_minutes = freq_mins,
      n_intraday_returns = length(ret_vec),
      lag1_return_acf = lag1_acf,
      lag1_to_lag5_acf_sum = lag1_to_lag5_acf_sum,
      ljung_box_pvalue_lag10 = ljung_p,
      mean_daily_rv = mean(rv_vec, na.rm = TRUE),
      mean_daily_bp = mean(bp_vec, na.rm = TRUE),
      mean_daily_rk = mean(rk_vec, na.rm = TRUE),
      mean_daily_ts = mean(ts_tbl$value, na.rm = TRUE),
      mean_daily_rts = mean(rts_tbl$value, na.rm = TRUE),
      rv_minus_rk = mean(rv_vec - rk_vec, na.rm = TRUE),
      rv_minus_rts = mean(rv_vec - rts_tbl$value, na.rm = TRUE)
    )
  }), fill = TRUE)
}), fill = TRUE)

setorder(microstructure_period_extremes, period_id, sample_minutes)
microstructure_period_extremes[]
```

```{r microstructure-noise-diagnostics-extreme-periods-plots, fig.width=11, fig.height=12}
if (nrow(microstructure_period_extremes) == 0L) {
  stop("No microstructure diagnostics available for selected periods.")
}

plot_tbl <- microstructure_period_extremes[is.finite(sample_minutes)]
period_ids_plot <- unique(plot_tbl$period_id)
palette_cols <- c("#D95F02", "#1B9E77", "#7570B3", "#E7298A")
period_colors <- setNames(
  palette_cols[seq_len(length(period_ids_plot))],
  period_ids_plot
)

plot_metric_by_period <- function(metric_col, ylab_txt, main_txt) {
  y_values <- plot_tbl[[metric_col]]
  y_finite <- y_values[is.finite(y_values)]
  y_lim <- if (length(y_finite) == 0L) {
    c(0, 1)
  } else {
    range(y_finite)
  }
  if (abs(y_lim[2] - y_lim[1]) < .Machine$double.eps) {
    y_lim <- y_lim + c(-0.1, 0.1)
  }

  plot(
    NA,
    xlim = range(sample_freq_grid),
    ylim = y_lim,
    xlab = "Sampling Frequency (minutes)",
    ylab = ylab_txt,
    main = main_txt
  )

  for (pid in period_ids_plot) {
    sub <- plot_tbl[period_id == pid]
    sub <- sub[is.finite(sub[[metric_col]])]
    if (nrow(sub) > 0L) {
      lines(
        sub$sample_minutes,
        sub[[metric_col]],
        type = "b",
        pch = 19,
        col = period_colors[[pid]]
      )
    }
  }

  legend(
    "topright",
    legend = period_ids_plot,
    col = unname(period_colors[period_ids_plot]),
    pch = 19,
    lty = 1,
    bty = "n"
  )
}

op <- par(no.readonly = TRUE)
par(mfrow = c(4, 1), mar = c(4, 4, 3.2, 1))

plot_metric_by_period(
  metric_col = "mean_daily_rv",
  ylab_txt = "Mean Daily RV",
  main_txt = "Volatility Signature: Highest vs Lowest Rejection Period"
)

plot_metric_by_period(
  metric_col = "lag1_return_acf",
  ylab_txt = "Lag-1 Return ACF",
  main_txt = "Lag-1 ACF: Highest vs Lowest Rejection Period"
)
abline(h = 0, lty = 2, col = "gray40")

plot_metric_by_period(
  metric_col = "lag1_to_lag5_acf_sum",
  ylab_txt = "Sum ACF (Lags 1-5)",
  main_txt = "Sum of Return ACF Lags 1-5: Highest vs Lowest Rejection Period"
)
abline(h = 0, lty = 2, col = "gray40")

plot_metric_by_period(
  metric_col = "rv_minus_rk",
  ylab_txt = "Mean(RV - RK)",
  main_txt = "Noise Gap (RV - RK): Highest vs Lowest Rejection Period"
)
abline(h = 0, lty = 2, col = "gray40")

par(op)
```

```{r microstructure-noise-diagnosis-gap-table}
if (length(selected_period_ids) >= 2L) {
  high_id <- highest_reject_period$period_id
  low_id <- lowest_reject_period$period_id

  high_tbl <- microstructure_period_extremes[
    period_id == high_id,
    .(
      sample_minutes,
      high_lag1_return_acf = lag1_return_acf,
      high_lag1_to_lag5_acf_sum = lag1_to_lag5_acf_sum,
      high_rv_minus_rk = rv_minus_rk,
      high_mean_daily_rv = mean_daily_rv
    )
  ]
  low_tbl <- microstructure_period_extremes[
    period_id == low_id,
    .(
      sample_minutes,
      low_lag1_return_acf = lag1_return_acf,
      low_lag1_to_lag5_acf_sum = lag1_to_lag5_acf_sum,
      low_rv_minus_rk = rv_minus_rk,
      low_mean_daily_rv = mean_daily_rv
    )
  ]

  microstructure_noise_gap <- merge(
    high_tbl,
    low_tbl,
    by = "sample_minutes",
    all = TRUE
  )
  setorder(microstructure_noise_gap, sample_minutes)
  microstructure_noise_gap[, `:=`(
    lag1_acf_gap_high_minus_low = high_lag1_return_acf - low_lag1_return_acf,
    lag1_to_lag5_acf_sum_gap_high_minus_low = high_lag1_to_lag5_acf_sum - low_lag1_to_lag5_acf_sum,
    rv_minus_rk_gap_high_minus_low = high_rv_minus_rk - low_rv_minus_rk,
    mean_daily_rv_gap_high_minus_low = high_mean_daily_rv - low_mean_daily_rv
  )]

  microstructure_noise_gap[]
} else {
  data.table(message = "Only one unique period found; two-period gap table is unavailable.")
}
```

```{r rejection-rate-by-period-plot, fig.width=10, fig.height=4}
plot_data <- rejection_rate_by_period[is.finite(rejection_rate)]

if (nrow(plot_data) == 0L) {
  stop("No finite rejection rates are available for plotting.")
}

plot(
  x = seq_len(nrow(plot_data)),
  y = plot_data$rejection_rate,
  type = "b",
  pch = 16,
  col = "#2C7FB8",
  ylim = c(0, 1),
  xaxt = "n",
  xlab = paste0("Period (", period_unit, ")"),
  ylab = "Rejection rate",
  main = "Rejection Rate by Period"
)
axis(1, at = seq_len(nrow(plot_data)), labels = plot_data$period_id, las = 2, cex.axis = 0.8)
abline(h = seq(0, 1, by = 0.1), col = "grey90", lty = "dotted")
```

```{r largest-ztest-day-by-period}
largest_ztest_by_period <- periodized_daily_tests[
  is.finite(ztest),
  .SD[which.max(ztest)],
  by = period_id
][
  ,
  .(
    period_id,
    period_start,
    period_end,
    DATE,
    ztest,
    pvalue,
    reject_null
  )
]

setorder(largest_ztest_by_period, period_start, DATE)
largest_ztest_by_period[]
```

```{r top-k-largest-ztest-days}
top_k_days <- 10L

top_k_ztest_days <- daily_tests[
  is.finite(ztest),
  .(DATE, ztest, pvalue, reject_null)
][order(-ztest, DATE)]

top_k_ztest_days <- head(top_k_ztest_days, top_k_days)
top_k_ztest_days <- merge(
  top_k_ztest_days,
  unique(periodized_daily_tests[, .(DATE, period_id, period_start, period_end)]),
  by = "DATE",
  all.x = TRUE
)
setcolorder(
  top_k_ztest_days,
  c("period_id", "period_start", "period_end", "DATE", "ztest", "pvalue", "reject_null")
)
setorder(top_k_ztest_days, -ztest, DATE)
top_k_ztest_days[]
```

```{r period-day-price-plot}
visualized_day <- as.Date("2026-01-19")

if (nrow(largest_ztest_by_period) == 0) {
  stop("No finite ztest value is available for plotting.")
}

if (is.null(period_to_plot)) {
  period_to_plot <- largest_ztest_by_period$period_id[1]
}
if (!(period_to_plot %in% largest_ztest_by_period$period_id)) {
  stop("period_to_plot not found in available periods: ", period_to_plot)
}

if (is.null(visualized_day)) {
  day_to_plot <- largest_ztest_by_period[period_id == period_to_plot, DATE][1]
} else {
  day_to_plot <- as.Date(visualized_day)
}

plot_price_data <- resample_prices(
  pData,
  freq_mins = plot_sample_minutes,
  tz = "UTC"
)

plot_prices <- plot_price_data[as.Date(DT, tz = "UTC") == day_to_plot]
if (nrow(plot_prices) == 0) {
  stop("No price data found for selected day: ", as.character(day_to_plot))
}

day_test <- periodized_daily_tests[DATE == day_to_plot][1]

plot(
  plot_prices$DT,
  plot_prices$PRICE,
  type = "l",
  col = "black",
  lwd = 1.2,
  xlab = "Time (UTC)",
  ylab = "Price",
  main = paste0(
    "BTC Prices on ", day_to_plot,
    " (", plot_sample_minutes, "-minute sampling, ",
    ", z=",
    sprintf("%.2f", day_test$ztest), ")"
  )
)

legend(
  "topleft",
  legend = "Price",
  col = "black",
  lty = 1,
  bty = "n"
)
```

```{r period-day-price-zoom-plot}
# Optional zoom window (UTC) within the selected day_to_plot
zoom_start_time <- "01:30:00"
zoom_end_time <- "02:30:00"

zoom_start_dt <- as.POSIXct(
  paste(as.character(day_to_plot), zoom_start_time),
  tz = "UTC"
)
zoom_end_dt <- as.POSIXct(
  paste(as.character(day_to_plot), zoom_end_time),
  tz = "UTC"
)

plot_prices_zoom <- plot_price_data[DT >= zoom_start_dt & DT <= zoom_end_dt]
if (nrow(plot_prices_zoom) == 0) {
  stop(
    "No price data in selected zoom window: ",
    format(zoom_start_dt, "%Y-%m-%d %H:%M:%S"),
    " to ",
    format(zoom_end_dt, "%Y-%m-%d %H:%M:%S"),
    " (UTC)"
  )
}

plot(
  plot_prices_zoom$DT,
  plot_prices_zoom$PRICE,
  type = "l",
  col = "black",
  lwd = 1.2,
  xlab = "Time (UTC)",
  ylab = "Price",
  main = paste0(
    "Zoomed BTC Prices on ", day_to_plot, " (",
    zoom_start_time, " - ", zoom_end_time, " UTC)"
  )
)

legend(
  "topleft",
  legend = "Price",
  col = "black",
  lty = 1,
  bty = "n"
)
```

# 10) Intraday Jump Test for a Specific Day (with Previous-Day Warmup)

```{r intraday-specific-day-controls}
# Keep these settings consistent with btc_jump_tests.Rmd::intraday-jump-test
intraday_lookback_period <- 90L
intraday_align_by <- "minutes" 
intraday_align_period <- 1L * 5
intraday_market_open <- "00:00:00"
intraday_market_close <- "23:59:59"
intraday_tz <- "UTC"

# Choose the day to test. If NULL, fallback to day_to_plot from section 9.
intraday_test_day <- as.Date("2025-12-26")

# Include the previous day so rolling volatility estimates at the start of
# intraday_test_day have warmup observations.
intraday_warmup_days <- 30L

if (is.null(intraday_test_day)) {
  intraday_test_day <- as.Date(day_to_plot)
}
intraday_test_day <- as.Date(intraday_test_day)

intraday_window_start <- intraday_test_day - intraday_warmup_days
intraday_window_end <- intraday_test_day

intraday_start_dt <- as.POSIXct(
  paste(as.character(intraday_window_start), "00:00:00"),
  tz = intraday_tz
)
intraday_end_dt <- as.POSIXct(
  paste(as.character(intraday_window_end), "23:59:59"),
  tz = intraday_tz
)

pData_intraday_test <- pData[
  DT >= intraday_start_dt & DT <= intraday_end_dt
]

if (nrow(pData_intraday_test) == 0) {
  stop(
    "No data found in intraday test window: ",
    format(intraday_start_dt, "%Y-%m-%d %H:%M:%S"),
    " to ",
    format(intraday_end_dt, "%Y-%m-%d %H:%M:%S"),
    " (UTC)"
  )
}

cat("Intraday target day:", as.character(intraday_test_day), "\n")
cat("Warmup days included:", intraday_warmup_days, "\n")
cat(
  "Intraday estimation window:",
  format(intraday_start_dt, "%Y-%m-%d %H:%M:%S"),
  "to",
  format(intraday_end_dt, "%Y-%m-%d %H:%M:%S"),
  "UTC\n"
)
cat("Rows in intraday test window:", nrow(pData_intraday_test), "\n")
```

```{r intraday-specific-day-test}
intraday_specific <- intradayJumpTest(
  pData = pData_intraday_test,
  alignBy = intraday_align_by,
  alignPeriod = intraday_align_period,
  marketOpen = intraday_market_open,
  marketClose = intraday_market_close,
  tz = intraday_tz,
)

if (!inherits(intraday_specific, "intradayJumpTest")) {
  stop("Expected intraday_specific to be class intradayJumpTest.")
}
if (!xts::is.xts(intraday_specific$ztest)) {
  stop("Expected intraday_specific$ztest to be xts.")
}

intraday_critical <- intraday_specific$criticalValue
if (is.null(intraday_critical)) {
  intraday_critical <- intraday_specific$critical.value
}
intraday_critical <- as.numeric(intraday_critical)[1]

intraday_specific_df <- data.table(
  DT = as.POSIXct(zoo::index(intraday_specific$ztest), tz = intraday_tz),
  ztest = as.numeric(intraday_specific$ztest)
)
intraday_specific_df[, criticalValue := intraday_critical]
intraday_specific_df[, jump_flag := !is.na(ztest) & is.finite(ztest) & (abs(ztest) > criticalValue)]
intraday_specific_df[, DATE := as.Date(DT, tz = intraday_tz)]

intraday_specific_day <- intraday_specific_df[DATE == intraday_test_day]
intraday_specific_jumps <- intraday_specific_day[jump_flag == TRUE]

cat("Intraday rows on target day:", nrow(intraday_specific_day), "\n")
cat("Detected intraday jumps on target day:", nrow(intraday_specific_jumps), "\n")

intraday_specific_jumps
```

```{r}
plot(intraday_specific$ztest)
```

```{r intraday-specific-day-plot}
intraday_plot_data <- resample_prices(
  pData_intraday_test,
  freq_mins = intraday_align_period,
  tz = intraday_tz
)[as.Date(DT, tz = intraday_tz) == intraday_test_day]

if (nrow(intraday_plot_data) == 0) {
  stop("No intraday plot data found for target day: ", as.character(intraday_test_day))
}

jump_plot_data <- merge(
  intraday_plot_data,
  intraday_specific_jumps[, .(DT, ztest)],
  by = "DT",
  all = FALSE
)

plot(
  intraday_plot_data$DT,
  intraday_plot_data$PRICE,
  type = "l",
  col = "black",
  lwd = 1.2,
  xlab = "Time (UTC)",
  ylab = "Price",
  main = paste0(
    "Intraday Test on ", intraday_test_day,
    " (", intraday_align_period, "-minute, lookBack=", intraday_lookback_period,
    ", warmup days=", intraday_warmup_days, ")"
  )
)

if (nrow(jump_plot_data) > 0) {
  points(
    jump_plot_data$DT,
    jump_plot_data$PRICE,
    pch = 19,
    col = "red",
    cex = 0.9
  )
  text(
    jump_plot_data$DT,
    jump_plot_data$PRICE,
    labels = sprintf("z=%.2f", jump_plot_data$ztest),
    pos = 4,
    cex = 0.7,
    col = "red"
  )
}

legend(
  "topleft",
  legend = c("Price", "Detected intraday jump"),
  col = c("black", "red"),
  lty = c(1, NA),
  pch = c(NA, 19),
  bty = "n"
)
```

```{r intraday-specific-day-zoom-plot}
# Optional zoom window (UTC) within intraday_test_day
intraday_zoom_start_time <- "07:00:00"
intraday_zoom_end_time <- "09:00:00"

intraday_zoom_start_dt <- as.POSIXct(
  paste(as.character(intraday_test_day), intraday_zoom_start_time),
  tz = intraday_tz
)
intraday_zoom_end_dt <- as.POSIXct(
  paste(as.character(intraday_test_day), intraday_zoom_end_time),
  tz = intraday_tz
)

intraday_plot_data_zoom <- intraday_plot_data[
  DT >= intraday_zoom_start_dt & DT <= intraday_zoom_end_dt
]
if (nrow(intraday_plot_data_zoom) == 0) {
  stop(
    "No intraday data in selected zoom window: ",
    format(intraday_zoom_start_dt, "%Y-%m-%d %H:%M:%S"),
    " to ",
    format(intraday_zoom_end_dt, "%Y-%m-%d %H:%M:%S"),
    " (UTC)"
  )
}

intraday_specific_jumps_zoom <- intraday_specific_jumps[
  DT >= intraday_zoom_start_dt & DT <= intraday_zoom_end_dt,
  .(DT, ztest)
]
jump_plot_data_zoom <- merge(
  intraday_plot_data_zoom,
  intraday_specific_jumps_zoom,
  by = "DT",
  all = FALSE
)

plot(
  intraday_plot_data_zoom$DT,
  intraday_plot_data_zoom$PRICE,
  type = "l",
  col = "black",
  lwd = 1.2,
  xlab = "Time (UTC)",
  ylab = "Price",
  main = paste0(
    "Zoomed Intraday Prices on ", intraday_test_day, " (",
    intraday_zoom_start_time, " - ", intraday_zoom_end_time, " UTC)"
  )
)

if (nrow(jump_plot_data_zoom) > 0) {
  points(
    jump_plot_data_zoom$DT,
    jump_plot_data_zoom$PRICE,
    pch = 19,
    col = "red",
    cex = 0.9
  )
  text(
    jump_plot_data_zoom$DT,
    jump_plot_data_zoom$PRICE,
    labels = sprintf("z=%.2f", jump_plot_data_zoom$ztest),
    pos = 4,
    cex = 0.7,
    col = "red"
  )
}

legend(
  "topleft",
  legend = c("Price", "Detected intraday jump"),
  col = c("black", "red"),
  lty = c(1, NA),
  pch = c(NA, 19),
  bty = "n"
)
```

# 11) April vs May Jump Dynamic Changes

```{r april-may-target-periods}
if (!exists("daily_tests") || nrow(daily_tests) == 0L) {
  stop("Section 11 needs `daily_tests` from section 9.")
}

monthly_daily_tests <- copy(daily_tests)
monthly_daily_tests[, period_id := format(DATE, "%Y-%m")]
monthly_daily_tests[, month_num := as.integer(format(DATE, "%m"))]

monthly_rejection_rank <- monthly_daily_tests[
  ,
  .(
    period_start = min(DATE),
    period_end = max(DATE),
    n_days = .N,
    reject_days = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE),
    mean_ztest = mean(ztest, na.rm = TRUE),
    mean_abs_ztest = mean(abs(ztest), na.rm = TRUE),
    median_pvalue = median(pvalue, na.rm = TRUE)
  ),
  by = .(period_id, month_num)
]
setorder(monthly_rejection_rank, period_start)
monthly_rejection_rank[, global_rank_high_to_low := rank(-rejection_rate, ties.method = "min")]
monthly_rejection_rank[, global_rank_low_to_high := rank(rejection_rate, ties.method = "min")]

april_period_id <- monthly_rejection_rank[
  month_num == 4L
][order(-rejection_rate, period_start), period_id][1]
may_period_id <- monthly_rejection_rank[
  month_num == 5L
][order(rejection_rate, period_start), period_id][1]

if (is.na(april_period_id) || is.na(may_period_id)) {
  stop("Could not locate both April and May periods in the current sample.")
}

april_may_period_ids <- c(april_period_id, may_period_id)
april_may_monthly_context <- monthly_rejection_rank[period_id %in% april_may_period_ids]
april_may_monthly_context[, role := ifelse(
  period_id == april_period_id,
  "April reference (highest among Aprils)",
  "May reference (lowest among Mays)"
)]
setorder(april_may_monthly_context, -rejection_rate, period_start)

monthly_rejection_rank[]
april_may_monthly_context[]
```

```{r april-may-jump-mechanism}
april_may_sampled_prices <- resample_prices(
  pData,
  freq_mins = fixed_sample_minutes,
  tz = "UTC"
)
april_may_sampled_prices[, DATE := as.Date(DT, tz = "UTC")]
april_may_sampled_prices[, period_id := format(DATE, "%Y-%m")]
april_may_sampled_prices <- april_may_sampled_prices[period_id %in% april_may_period_ids]

if (nrow(april_may_sampled_prices) == 0L) {
  stop("No sampled prices for selected April/May periods.")
}

april_may_daily_hf <- rbindlist(lapply(april_may_period_ids, function(pid) {
  px_period <- april_may_sampled_prices[period_id == pid]
  px_xts <- xts(px_period$PRICE, order.by = px_period$DT, tzone = "UTC")

  rv_xts <- rCov(px_xts, makeReturns = TRUE)
  bp_xts <- rBPCov(px_xts, makeReturns = TRUE)
  medrv_xts <- rMedRVar(px_xts, makeReturns = TRUE)
  minrv_xts <- rMinRVar(px_xts, makeReturns = TRUE)
  quar_xts <- rQuar(px_xts, makeReturns = TRUE)
  tpquar_xts <- rTPQuar(px_xts, makeReturns = TRUE)

  out <- data.table(
    DATE = as.Date(zoo::index(rv_xts), tz = "UTC"),
    period_id = pid,
    rv = to_numeric_xts(rv_xts),
    bpv = to_numeric_xts(bp_xts),
    medrv = to_numeric_xts(medrv_xts),
    minrv = to_numeric_xts(minrv_xts),
    quar = to_numeric_xts(quar_xts),
    tpquar = to_numeric_xts(tpquar_xts)
  )

  out[, jump_gap := pmax(rv - bpv, 0)]
  out[, jump_share_rv := ifelse(rv > 0, jump_gap / rv, NA_real_)]
  out[, bpv_share_rv := ifelse(rv > 0, bpv / rv, NA_real_)]
  out[, jump_signal_scaled := jump_gap / sqrt(pmax(tpquar, .Machine$double.eps))]
  out
}), fill = TRUE)

april_may_daily_returns <- copy(april_may_sampled_prices)
april_may_daily_returns[, ret := log(PRICE) - shift(log(PRICE), type = "lag"), by = .(period_id, DATE)]
april_may_daily_tail <- april_may_daily_returns[is.finite(ret), .(
  max_abs_ret = max(abs(ret), na.rm = TRUE),
  q99_abs_ret = as.numeric(quantile(abs(ret), 0.99, na.rm = TRUE))
), by = .(period_id, DATE)]

april_may_daily_features <- merge(
  monthly_daily_tests[period_id %in% april_may_period_ids, .(DATE, period_id, reject_null, ztest, pvalue)],
  april_may_daily_hf,
  by = c("DATE", "period_id"),
  all.x = TRUE
)
april_may_daily_features <- merge(
  april_may_daily_features,
  april_may_daily_tail,
  by = c("DATE", "period_id"),
  all.x = TRUE
)
april_may_daily_features <- april_may_daily_features[is.finite(rv) & is.finite(bpv)]

jump_share_q75 <- quantile(april_may_daily_features$jump_share_rv, 0.75, na.rm = TRUE)

april_may_mechanism_summary <- april_may_daily_features[
  ,
  .(
    n_days = .N,
    reject_days = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE),
    mean_abs_ztest = mean(abs(ztest), na.rm = TRUE),
    median_pvalue = median(pvalue, na.rm = TRUE),
    mean_rv = mean(rv, na.rm = TRUE),
    mean_bpv_share_rv = mean(bpv_share_rv, na.rm = TRUE),
    mean_jump_share_rv = mean(jump_share_rv, na.rm = TRUE),
    high_jump_day_share = mean(jump_share_rv > jump_share_q75, na.rm = TRUE),
    mean_jump_signal_scaled = mean(jump_signal_scaled, na.rm = TRUE),
    p90_jump_signal_scaled = as.numeric(quantile(jump_signal_scaled, 0.90, na.rm = TRUE)),
    mean_tpquar = mean(tpquar, na.rm = TRUE),
    mean_max_abs_ret = mean(max_abs_ret, na.rm = TRUE),
    mean_q99_abs_ret = mean(q99_abs_ret, na.rm = TRUE)
  ),
  by = period_id
]
setorder(april_may_mechanism_summary, -rejection_rate)

april_may_mechanism_summary[]
```

```{r april-may-frequency-confidence-sensitivity}
april_may_freq_conf <- copy(daily_details_all)
april_may_freq_conf[, period_id := format(DATE, "%Y-%m")]
april_may_freq_conf <- april_may_freq_conf[period_id %in% april_may_period_ids]

april_may_freq_conf_summary <- april_may_freq_conf[
  ,
  .(
    n_days = .N,
    reject_days = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE),
    mean_abs_ztest = mean(abs(ztest), na.rm = TRUE)
  ),
  by = .(period_id, sample_minutes, confidence_level)
]
setorder(april_may_freq_conf_summary, confidence_level, sample_minutes, period_id)

april_may_freq_conf_wide <- dcast(
  april_may_freq_conf_summary,
  sample_minutes + confidence_level ~ period_id,
  value.var = "rejection_rate"
)

if (all(april_may_period_ids %in% names(april_may_freq_conf_wide))) {
  april_may_freq_conf_wide[, april_minus_may := get(april_period_id) - get(may_period_id)]
}

april_may_freq_conf_summary[]
april_may_freq_conf_wide[]
```

```{r april-may-bns-robustness}
run_daily_bns_with_estimators <- function(sampled_pdata, confidence_level, iv_est, iq_est) {
  alpha_quantile <- 1 - (1 - confidence_level) / 2
  bns_obj <- BNSjumpTest(
    rData = sampled_pdata,
    IVestimator = iv_est,
    IQestimator = iq_est,
    type = "ratio",
    max = TRUE,
    makeReturns = TRUE,
    alpha = alpha_quantile
  )
  extract_daily_bns_summary(
    bns_obj,
    confidence_level = confidence_level,
    tz = "UTC"
  )
}

bns_robust_variants <- list(
  list(label = "BNS_BV_TP", iv = "BV", iq = "TP"),
  list(label = "BNS_MedRV_TPQuar", iv = "rMedRVar", iq = "rTPQuar"),
  list(label = "BNS_MinRV_TPQuar", iv = "rMinRVar", iq = "rTPQuar")
)

april_may_robust_results <- rbindlist(lapply(bns_robust_variants, function(v) {
  daily <- run_daily_bns_with_estimators(
    sampled_pdata = april_may_sampled_prices[, .(DT, PRICE)],
    confidence_level = fixed_confidence_level,
    iv_est = v$iv,
    iq_est = v$iq
  )
  daily[, period_id := format(DATE, "%Y-%m")]
  daily <- daily[period_id %in% april_may_period_ids]
  daily[, variant := v$label]
  daily
}), fill = TRUE)

april_may_robust_summary <- april_may_robust_results[
  ,
  .(
    n_days = .N,
    reject_days = sum(reject_null, na.rm = TRUE),
    rejection_rate = mean(reject_null, na.rm = TRUE),
    mean_abs_ztest = mean(abs(ztest), na.rm = TRUE),
    median_pvalue = median(pvalue, na.rm = TRUE)
  ),
  by = .(variant, period_id)
]
setorder(april_may_robust_summary, variant, -rejection_rate)

april_may_robust_summary[]
```

```{r april-may-diagnostics-plot, fig.width=12, fig.height=8}
if (nrow(april_may_freq_conf_summary) == 0L || nrow(april_may_daily_features) == 0L) {
  stop("No April-May diagnostics data to plot.")
}

plot_95 <- april_may_freq_conf_summary[abs(confidence_level - fixed_confidence_level) < 1e-10]
plot_95_ids <- unique(plot_95$period_id)
cols <- setNames(c("#D95F02", "#1B9E77", "#7570B3"), plot_95_ids)

op <- par(no.readonly = TRUE)
par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

plot(
  NA,
  xlim = range(sample_freq_grid),
  ylim = c(0, 1),
  xlab = "Sampling Frequency (minutes)",
  ylab = "Rejection rate",
  main = paste0("Rejection by Sampling Frequency (", fixed_confidence_level * 100, "% CI)")
)
for (pid in plot_95_ids) {
  sub <- plot_95[period_id == pid]
  lines(sub$sample_minutes, sub$rejection_rate, type = "b", pch = 19, col = cols[[pid]])
}
legend("topright", legend = plot_95_ids, col = unname(cols[plot_95_ids]), pch = 19, lty = 1, bty = "n")

boxplot(
  jump_share_rv ~ period_id,
  data = april_may_daily_features,
  col = c("#FDD0A2", "#A1D99B"),
  ylab = "Jump share of RV: max(RV - BPV, 0) / RV",
  xlab = "Period",
  main = "Daily Jump Share Distribution"
)

boxplot(
  jump_signal_scaled ~ period_id,
  data = april_may_daily_features,
  col = c("#FEE391", "#9ECAE1"),
  ylab = "(RV - BPV)+ / sqrt(TPQuar)",
  xlab = "Period",
  main = "Scaled Jump Signal"
)

boxplot(
  max_abs_ret ~ period_id,
  data = april_may_daily_features,
  col = c("#FC9272", "#74C476"),
  ylab = "Max |30m log return| per day",
  xlab = "Period",
  main = "Intraday Tail Size"
)

par(op)
```

```{r april-may-explanation-table}
april_row <- april_may_mechanism_summary[period_id == april_period_id][1]
may_row <- april_may_mechanism_summary[period_id == may_period_id][1]

if (nrow(april_row) == 0L || nrow(may_row) == 0L) {
  stop("Could not build April-May explanation table.")
}

april_may_explanation <- data.table(
  comparison = paste0(april_period_id, " minus ", may_period_id),
  rejection_rate_gap = april_row$rejection_rate - may_row$rejection_rate,
  mean_abs_ztest_gap = april_row$mean_abs_ztest - may_row$mean_abs_ztest,
  mean_jump_share_gap = april_row$mean_jump_share_rv - may_row$mean_jump_share_rv,
  high_jump_day_share_gap = april_row$high_jump_day_share - may_row$high_jump_day_share,
  mean_scaled_jump_signal_gap = april_row$mean_jump_signal_scaled - may_row$mean_jump_signal_scaled,
  mean_max_abs_ret_gap = april_row$mean_max_abs_ret - may_row$mean_max_abs_ret,
  mean_tpquar_gap = april_row$mean_tpquar - may_row$mean_tpquar
)

april_may_explanation[]
```


