---
title: "OKX HF EDA: Leverage Effect and Microstructure Noise (2601 + 2602)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## 1) Install and Load Packages

```{r packages}
required_pkgs <- c("data.table", "highfrequency", "xts", "zoo", "ggplot2")

for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
}

library(data.table)
library(highfrequency)
library(xts)
library(zoo)
library(ggplot2)
```

## 2) Load and Concatenate Data

```{r load-data}
input_paths <- c(
  "target/eth_2601_1m.csv",
  "target/eth_2602_1m.csv"
)

missing_paths <- input_paths[!file.exists(input_paths)]
if (length(missing_paths) > 0) {
  stop("Missing input file(s): ", paste(missing_paths, collapse = ", "))
}

hf_list <- lapply(input_paths, fread)
hf <- rbindlist(hf_list, use.names = TRUE, fill = TRUE, idcol = "source_id")
hf[, source_file := input_paths[source_id]]

required_cols <- c("DT", "PRICE", "BID", "OFR", "MIDQUOTE")
if (!all(required_cols %in% names(hf))) {
  stop(
    "Input data must contain columns: ",
    paste(required_cols, collapse = ", ")
  )
}

numeric_cols <- c("PRICE", "BID", "OFR", "MIDQUOTE", "SIZE", "BIDSIZ", "OFRSIZ")
available_numeric_cols <- intersect(numeric_cols, names(hf))

hf[, DT := as.POSIXct(DT, tz = "UTC")]
hf[, (available_numeric_cols) := lapply(.SD, as.numeric), .SDcols = available_numeric_cols]

hf <- hf[
  !is.na(DT) &
    is.finite(PRICE) & PRICE > 0 &
    is.finite(BID) & BID > 0 &
    is.finite(OFR) & OFR > 0 &
    OFR >= BID
]

setorder(hf, DT)
hf[, DATE := as.Date(DT, tz = "UTC")]

cat("Rows after cleaning:", nrow(hf), "\n")
cat("Date range:", as.character(min(hf$DT)), "to", as.character(max(hf$DT)), "\n")
cat("Number of UTC dates:", uniqueN(hf$DATE), "\n")
cat("Input files:", paste(unique(hf$source_file), collapse = ", "), "\n")

head(hf)
```

## 3) Quick Data Sanity Checks

```{r sanity-checks}
daily_counts <- hf[, .N, by = DATE][order(DATE)]
daily_spread <- hf[, .(
  mean_spread = mean(OFR - BID, na.rm = TRUE),
  median_spread = median(OFR - BID, na.rm = TRUE),
  mean_rel_spread_bps = mean((OFR - BID) / MIDQUOTE * 1e4, na.rm = TRUE)
), by = DATE][order(DATE)]

summary(daily_counts$N)
summary(daily_spread$mean_rel_spread_bps)
head(daily_counts)
head(daily_spread)
```

## 4) Helper Functions

```{r helper-functions}
sample_freqs <- c(1L, 2L, 5L, 10L, 15L, 30L, 60L)

resample_hf <- function(hf_data, freq_mins, tz = "UTC") {
  dt <- copy(hf_data[, .(DT, PRICE, BID, OFR, MIDQUOTE)])
  if (freq_mins > 1L) {
    dt <- as.data.table(
      aggregatePrice(
        dt,
        alignBy = "minutes",
        alignPeriod = freq_mins,
        marketOpen = "00:00:00",
        marketClose = "23:59:59",
        tz = tz
      )
    )
  }

  dt <- dt[
    is.finite(PRICE) & PRICE > 0 &
      is.finite(BID) & BID > 0 &
      is.finite(OFR) & OFR > 0 &
      is.finite(MIDQUOTE) & MIDQUOTE > 0 &
      OFR >= BID
  ]

  dt[, DATE := as.Date(DT, tz = tz)]
  valid_dates <- dt[, .N, by = DATE][N > 1L, DATE]
  dt <- dt[DATE %in% valid_dates]
  setorder(dt, DT)
  dt
}

make_intraday_returns <- function(sampled_dt, price_col = "PRICE") {
  out <- copy(sampled_dt)
  out[, RET := log(get(price_col)) - shift(log(get(price_col))), by = DATE]
  out[is.finite(RET)]
}

to_numeric_xts <- function(x) {
  as.numeric(zoo::coredata(x))
}

daywise_noise_robust <- function(sampled_dt, estimator = c("rTSCov", "rRTSCov"), tz = "UTC") {
  estimator <- match.arg(estimator)
  px_xts <- xts(sampled_dt$PRICE, order.by = sampled_dt$DT, tzone = tz)
  day_index <- as.Date(index(px_xts), tz = tz)
  split_rows <- split(seq_along(day_index), day_index)

  out <- rbindlist(lapply(names(split_rows), function(d) {
    rows <- split_rows[[d]]
    x_day <- px_xts[rows]
    n_obs <- NROW(x_day)

    if (n_obs < 20L) {
      return(data.table(DATE = as.Date(d), value = NA_real_, n_obs = n_obs))
    }

    # rTS/rRTS in highfrequency requires at least ~10*K observations.
    K_use <- max(2L, floor((n_obs - 1L) / 10L))
    K_use <- min(K_use, 50L)
    J_use <- 1L

    est_val <- tryCatch(
      {
        if (estimator == "rTSCov") {
          as.numeric(rTSCov(x_day, K = K_use, J = J_use))
        } else {
          as.numeric(rRTSCov(x_day, K = K_use, J = J_use))
        }
      },
      error = function(e) NA_real_
    )

    data.table(DATE = as.Date(d), value = est_val, n_obs = n_obs)
  }), fill = TRUE)

  setorder(out, DATE)
  out
}
```

## 5) Leverage Effect Diagnostics by Sampling Frequency

Leverage effect proxy used here: correlation between return at day `t` and volatility change from day `t` to `t+1`.

```{r leverage-diagnostics}
leverage_details <- lapply(sample_freqs, function(freq) {
  sampled <- resample_hf(hf, freq_mins = freq, tz = "UTC")
  rets <- make_intraday_returns(sampled, price_col = "PRICE")

  daily_panel <- sampled[, .(close_price = last(PRICE)), by = DATE]
  daily_rv <- rets[, .(rv = sum(RET^2, na.rm = TRUE)), by = DATE]
  daily <- merge(daily_panel, daily_rv, by = "DATE", all = FALSE)
  setorder(daily, DATE)

  daily[, close_ret := log(close_price) - shift(log(close_price))]
  daily[, rv_next := shift(rv, type = "lead")]
  daily[, vol_change_t1 := rv_next - rv]

  daily_model <- daily[
    is.finite(close_ret) &
      is.finite(vol_change_t1)
  ]

  if (nrow(daily_model) < 10L) {
    return(list(
      freq = freq,
      summary = data.table(
        sample_minutes = freq,
        n_days = nrow(daily_model),
        corr_ret_t_volchg_t1 = NA_real_,
        beta_close_ret = NA_real_,
        beta_pvalue = NA_real_,
        mean_volchg_if_ret_neg = NA_real_,
        mean_volchg_if_ret_pos = NA_real_,
        neg_pos_volchg_ratio = NA_real_
      ),
      daily = daily
    ))
  }

  fit <- lm(vol_change_t1 ~ close_ret, data = daily_model)
  fit_coef <- summary(fit)$coefficients

  mean_neg <- daily_model[close_ret < 0, mean(vol_change_t1, na.rm = TRUE)]
  mean_pos <- daily_model[close_ret >= 0, mean(vol_change_t1, na.rm = TRUE)]

  leverage_summary <- data.table(
    sample_minutes = freq,
    n_days = nrow(daily_model),
    corr_ret_t_volchg_t1 = cor(daily_model$close_ret, daily_model$vol_change_t1),
    beta_close_ret = unname(fit_coef["close_ret", "Estimate"]),
    beta_pvalue = unname(fit_coef["close_ret", "Pr(>|t|)"]),
    mean_volchg_if_ret_neg = mean_neg,
    mean_volchg_if_ret_pos = mean_pos,
    neg_pos_volchg_ratio = mean_neg / mean_pos
  )

  list(
    freq = freq,
    summary = leverage_summary,
    daily = daily
  )
})

leverage_summary <- rbindlist(lapply(leverage_details, `[[`, "summary"), fill = TRUE)
leverage_summary
```

```{r leverage-plot}
ggplot(leverage_summary, aes(x = sample_minutes, y = corr_ret_t_volchg_t1)) +
  geom_line(color = "#2A6F97") +
  geom_point(color = "#2A6F97", size = 2.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "Leverage Effect by Sampling Frequency",
    subtitle = "Correlation between close return (t) and volatility change (RV_t+1 - RV_t)",
    x = "Sampling frequency (minutes)",
    y = "Corr(ret_t, RV_t+1 - RV_t)"
  ) +
  theme_minimal(base_size = 12)
```

## 6) Microstructure Noise Diagnostics by Sampling Frequency

Diagnostics include:

- lag-1 return autocorrelation (bid-ask bounce proxy)
- volatility signature plot (mean daily RV vs frequency)
- robust-vs-naive realized variance comparison (`rKernelCov`, `rTSCov`, `rRTSCov`)
- average quoted spread and relative spread

```{r microstructure-diagnostics}
microstructure_details <- lapply(sample_freqs, function(freq) {
  sampled <- resample_hf(hf, freq_mins = freq, tz = "UTC")
  rets <- make_intraday_returns(sampled, price_col = "PRICE")
  ret_vec <- rets$RET

  lag1_acf <- if (length(ret_vec) > 1L) {
    cor(ret_vec[-1], ret_vec[-length(ret_vec)])
  } else {
    NA_real_
  }

  ljung_p <- if (length(ret_vec) > 20L) {
    Box.test(ret_vec, lag = 10L, type = "Ljung-Box")$p.value
  } else {
    NA_real_
  }

  spread_mean <- sampled[, mean(OFR - BID, na.rm = TRUE)]
  spread_rel_bps <- sampled[, mean((OFR - BID) / MIDQUOTE * 1e4, na.rm = TRUE)]

  px_xts <- xts(sampled$PRICE, order.by = sampled$DT, tzone = "UTC")
  rv_xts <- rCov(px_xts, makeReturns = TRUE)
  bp_xts <- rBPCov(px_xts, makeReturns = TRUE)
  rk_xts <- suppressWarnings(rKernelCov(px_xts, makeReturns = TRUE))

  rv_vec <- to_numeric_xts(rv_xts)
  bp_vec <- to_numeric_xts(bp_xts)
  rk_vec <- to_numeric_xts(rk_xts)
  rk_vec[rk_vec < 0] <- NA_real_

  ts_tbl <- daywise_noise_robust(sampled, estimator = "rTSCov", tz = "UTC")
  rts_tbl <- daywise_noise_robust(sampled, estimator = "rRTSCov", tz = "UTC")

  data.table(
    sample_minutes = freq,
    n_intraday_returns = length(ret_vec),
    lag1_return_acf = lag1_acf,
    ljung_box_pvalue_lag10 = ljung_p,
    mean_spread = spread_mean,
    mean_relative_spread_bps = spread_rel_bps,
    mean_daily_rv = mean(rv_vec, na.rm = TRUE),
    mean_daily_bp = mean(bp_vec, na.rm = TRUE),
    mean_daily_rk = mean(rk_vec, na.rm = TRUE),
    mean_daily_ts = mean(ts_tbl$value, na.rm = TRUE),
    mean_daily_rts = mean(rts_tbl$value, na.rm = TRUE),
    rv_minus_rk = mean(rv_vec - rk_vec, na.rm = TRUE),
    rv_minus_rts = mean(rv_vec - rts_tbl$value, na.rm = TRUE)
  )
})

microstructure_summary <- rbindlist(microstructure_details, fill = TRUE)
microstructure_summary
```

```{r microstructure-plots}
p_signature <- ggplot(microstructure_summary, aes(x = sample_minutes, y = mean_daily_rv)) +
  geom_line(color = "#CC5500") +
  geom_point(color = "#CC5500", size = 2.2) +
  labs(
    title = "Volatility Signature Plot",
    subtitle = "Mean daily realized variance decreases as sampling gets coarser",
    x = "Sampling frequency (minutes)",
    y = "Mean daily RV"
  ) +
  theme_minimal(base_size = 12)

p_acf <- ggplot(microstructure_summary, aes(x = sample_minutes, y = lag1_return_acf)) +
  geom_line(color = "#0B6E4F") +
  geom_point(color = "#0B6E4F", size = 2.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "Lag-1 Return Autocorrelation by Sampling Frequency",
    x = "Sampling frequency (minutes)",
    y = "Lag-1 autocorrelation"
  ) +
  theme_minimal(base_size = 12)

p_robust_gap <- ggplot(microstructure_summary, aes(x = sample_minutes, y = rv_minus_rk)) +
  geom_line(color = "#1F4E79") +
  geom_point(color = "#1F4E79", size = 2.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "Naive RV - Realized Kernel (Noise Gap)",
    x = "Sampling frequency (minutes)",
    y = "Mean(RV - RK)"
  ) +
  theme_minimal(base_size = 12)

p_signature
p_acf
p_robust_gap
```

## 7) One-Page Summary Tables

```{r summary-tables}
cat("Leverage-effect summary across sampling frequencies:\n")
leverage_summary

cat("\nMicrostructure-noise summary across sampling frequencies:\n")
microstructure_summary
```

## 8) Notes

- If leverage effect is present, `corr_ret_t_volchg_t1` and `beta_close_ret` are typically negative.
- If microstructure noise is present, very high-frequency returns often show stronger negative lag-1 autocorrelation and higher naive RV relative to noise-robust estimators.
- You can extend `sample_freqs` to include seconds-level frequencies if higher-resolution data are added later.
